---
title: "DATA 607 Assignment – Tidying and Transforming Data"
author: "Yohannes Deboch"
date: "March 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```


The chart above describes arrival delays for two airlines across five destinations. Your task is to:
(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above.
You’re encouraged to use a “wide” structure similar to how the information appears above, so
that you can practice tidying and transformations as described below.
(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy
and transform your data.
Read the data.
```{r}
dt <- read.csv("Numbersense.csv", stringsAsFactors = FALSE, na.strings = "")
head(dt)
```
Fill the airline columns

```{r}
dt <- dt %>%
    fill( X)
```
Remove empty rows
```{r}
dt <- na.omit(dt)
head(dt)
```
Correct airlines column names
```{r}
colnames(dt)[1:2] <- c("Airlines","Status")
head(dt)
```
Transform wide to long format
```{r}
dt <- dt %>%
    gather(Destination, count, 3:7)
```
Seperate status column

```{r}
dt <- dt %>%
    spread(Status, count)
```

(3) Perform analysis to compare the arrival delays for the two airlines.
```{r}
glimpse(dt)
```
Airlines, Destination are string variable, delayed and on time are integer.
```{r}
dt %>%
    group_by(Airlines)%>%
    summarise(mean = mean(delayed))
```
Mean delay of Alaska is 100 and AM WEST is 157. ALASKA airlines has less delay.
Add total column
```{r}
dt$total <- dt$delayed+dt$`on time`
dt$percent_del <- (dt$delayed/dt$total)*100
```
```{r}
ggplot(data = dt, aes(Destination, percent_del, fill = Airlines))+geom_bar(stat = "identity", position = "dodge")+theme_classic()+ggtitle("Percent Delayed")+ylab("Percent")+theme(plot.title = element_text(hjust = 0.5))
```
San Francisco had the most delayed followed by Seattle where Phoenix had the least delayed.
(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative
descriptions of your data cleanup work, analysis, and conclusions. Please include in your
homework submission:
 The URL to the .Rmd file in your Git Hub repository. and
  The URL for your rpubs.com web page.


